---
title: "CTD and Bottle Product Comparison"
output: html_document
---

## Setup 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# clear workspace for local development
rm(list = ls())

# assign relative path to directory
dir <- "/Users/jaxinewolfe/Documents/WHOI/NESLTER/ctd-bottles-transect/"
# set as working directory
setwd(dir)

# define R packages to require
libs <- c("tidyverse", "maps")
# load libraries
lapply(libs, require, character.only = TRUE)
```

## Utility Function: Read From API
```{r}

# Read in files from the REST API
# accepts the type of data (ctd metadata, bottle summary, chl, nut)
# accepts a vector of cruises to compile data from

read_from_api <- function(type, cruises) {
  # expand the cruises into a dataframe (avoids nested for loops)
  z <- expand.grid(cruises)
  
  # read in data based on the specified source
  if (type == "metadata") {
    urls <- paste0("https://nes-lter-data.whoi.edu/api/ctd/", z$Var1, "/metadata.csv")
    urls <- unlist(urls)
  }
  if (type == "summary") {
    urls <- paste0("https://nes-lter-data.whoi.edu/api/ctd/", z$Var1, "/bottle_summary.csv")
    urls <- unlist(urls)
  }
  if (type == "nutrient") {
    urls <- paste0("https://nes-lter-data.whoi.edu/api/nut/", z$Var1, ".csv")
    urls <- unlist(urls)
  }
  if (type == "chl") {
    urls <- paste0("https://nes-lter-data.whoi.edu/api/chl/", z$Var1, ".csv")
    urls <- unlist(urls)
  }
  if (type == "bottles") {
    urls <- paste0("https://nes-lter-data.whoi.edu/api/ctd/", z$Var1, "/bottles.csv")
    urls <- unlist(urls)
  }

  ## Cruise Compilation ##
  # case: more than one cruise given
  if (length(cruises) > 1) {
  # begin compilation  
  prev_cruise <- read_csv(urls[1])
  
  # loop through urls to compile cruise data into one file
  for (k in 2:length(urls)){
    # read in data per cruise
    next_cruise <- read_csv(urls[k])
    
    # bind the next cruise to the compiled cruise dataset
    all <- plyr::rbind.fill(prev_cruise, next_cruise)
    
    # if statment to reset the previous cruises until all cruises are read in
    if(k < length(urls)) {
      prev_cruise <- all
    }
  }
  return(all)

  # case: only one cruise is given
  } else {
   all <- read_csv(urls)
   return(all)
  }
}

```

## Utility Function: Find Paired Casts
```{r}

# packages required: tidyverse

find_paired_casts <- function(df, min.time) {
  
  cruiselist <- unique(df$cruise)
  # read in ctd data
  ctd_metadata <- read_from_api(type = "metadata", cruises = cruiselist)
  ctd_summary <- read_from_api(type = "summary", cruises = cruiselist)

  # define column to display paired cast
  df$paired_cast <- NA_character_
  
  for (j in 1:nrow(df)) {
    # store values for current row
    df_cruise <- df$cruise[j]
    df_cast <- df$cast[j]
    
    # isolate cast metadata and bottle summary
    cast_metadata <- ctd_metadata %>% filter(cruise == df_cruise & cast == df_cast)
    cast_summary <- ctd_summary %>% filter(cruise == df_cruise & cast == df_cast)
    
    # define the min and max time for the cast 
    df_date_max <- max(cast_summary$date)
    df_date_min <- min(cast_summary$date)
    
    # find the cast before from the bottle summary
    cast_before <- ctd_summary %>% 
      filter(cruise == df_cruise & cast == (df_cast-1))
    # find the cast after from the ctd metadata
    cast_after <- ctd_metadata %>% 
      filter(cruise == df_cruise & cast == (df_cast+1))
    
    # deterine the time difference
    time_before <- min(abs(as.numeric(difftime(df_date_min, cast_before$date, units = "mins"))),
                       na.rm = TRUE)
    time_after <- min(abs(as.numeric(difftime(df_date_max, cast_after$date, units = "mins"))),
                      na.rm = TRUE)
    
    # determine paired casts if they exist
    if (time_after > min.time & time_before > min.time) {
      j <- j + 1
      next
    }
    if (time_before < min.time) {
      # isolate cast metadata to extract station
      before_metadata <- ctd_metadata %>% filter(cruise == df_cruise & cast == (df_cast-1))
      if (isFALSE(cast_metadata$nearest_station == before_metadata$nearest_station)) {
        print(paste0("paired cast found for cruise ", df_cruise, 
                     " and cast ", df_cast, ", but nearest station did not match"))
        j <- j + 1
        next
      } else {
        df$paired_cast[j] <- df_cast - 1
      }
    }
    if (time_after < min.time) {
      if (isFALSE(cast_metadata$nearest_station == cast_after$nearest_station)) {
        print(paste0("paired cast found for cruise ", df_cruise, 
                     " and cast ", df_cast, ", but nearest station did not match"))
        j <- j + 1
        next
      } else {
        df$paired_cast[j] <- df_cast + 1
      }
    }
  }
  return(df)
}

# Example code
# test_df <- find_paired_casts(df, min.time = 50)

```


## Compare Cast Function

Cruises with up and downcast CTD profiles:
  • ar31b (bin-averaged)
  • ar32 (bin-averaged)
  • ar34b (not bin-averaged)
  • en644 (api doesn't display upcast)

CTD bottle data provides a list of niskins along with data from the CTD as it has been provided by the SeaBird CTD processing software. CTD cast provides profiles for each specified cast. Bottle summaries only have the upcast. The goal is to relate the datasets based on depth (depsm) or pressure (prdm) and compare across temperature (t090c) and salinity (sal00). 

Function to write: compare_cast(cruise, castnum) 
  • input desired cast and cruise number
  
  From bottle summary:
  • pulls appropriate CTD and truncates it with the upcast function 
  • pulls bottle summary for given cruise and truncates the dataset to the given castnum
  
  Comparison:
  • use nearest pressure to compare temp and salinity between the two tables
  • add columns that indicate logical and numerical difference in the values
  
```{r}

# function returns the upcast for a given CTD dataset
upcast <- function(dataset, depth_col){
  # find the max depth
  ind <- which.max(dataset[[depth_col]])
  # return upcast
  return(dataset[ind:nrow(dataset),])
}

# compare CTD profiles and bottle summary for a specified cruise and cast
compare_cast <- function(cruise, castnum) {
  
  # load bottle summary
  bottle <- read_from_api(type = "bottles", cruises = cruise) %>%
    filter(cast == castnum) %>%
    select(depsm, prdm, t090c, sal00)

  # load ctd profile
  # https://nes-lter-data.whoi.edu/api/ctd/en617/cast_12.csv
  ctd <- read_csv(paste0("https://nes-lter-data.whoi.edu/api/ctd/", cruise, "/cast_", castnum, ".csv")) %>%
    select(depsm, prdm, t090c, sal00)

  # isolate upcast from the ctd profile
  ctd_upcast <- upcast(ctd, depth_col = "depsm")
  
  # define fields for logical and numerical comparisons
  bottle$sal_check <- FALSE
  bottle$sal_diff <- NA_integer_
  bottle$temp_check <- FALSE
  bottle$temp_diff <- NA_integer_

  for (i in 1:nrow(bottle)) {
    # store values
    bottle_prdm <- bottle$prdm[i]
    # bottle_depsm <- bottle$depsm[i] # using prdm to merge for now
    bottle_sal <- bottle$sal00[i]
    bottle_temp <-bottle$t090c[i]
  
    # find the index of the nearest datetime
    ind <- which.min(abs(bottle_prdm - ctd_upcast$prdm))
    
    # store sal and temp at closest pressure
    ctd_sal <- ctd_upcast$sal00[ind]
    ctd_temp <- ctd_upcast$t090c[ind]
  
    # logical salinity comparison 
    if (round(bottle_sal,2) == round(ctd_sal,2)) {
      bottle$sal_check[i] <- TRUE
    } 
    # logical temp comparison 
    if (round(bottle_temp,2) == round(ctd_temp,2)) {
      bottle$temp_check[i] <- TRUE
    } 
    
    # store differences
    bottle$sal_diff[i] <- abs(bottle_sal - ctd_sal)
    bottle$temp_diff[i] <- abs(bottle_temp - ctd_temp)
  }
  return(bottle)
  }

```

## Compare CTD cast and bottles data

The data displayed will be from the bottle summary. Function adds columns displaying the logical and numerical check for temp and salinity. 
```{r}

# load metadata
# metadata_ar31b <- read_from_api(type = "metadata", cruises = "ar31b")

compare_ar31b_6 <- compare_cast("ar31b", castnum = 6)

```
